\mutagen: Coverage-Guided, Property-Based Testing using Exhaustive Type-Preserving Mutations

### 2021-08-26 papers/icse22/main.tex
Heavy compression. Paper now fits!

%
% FuzzChick Limitations
%
\subsubsection{Limitations of CGPT}

@@

%% At this point, the reader might consider:
%% %
%% what about large enumeration types like integers or characters?
%% %
%% Does \mutagen transform them exaustively too?
%% %
%% \emph{Certainly not.}
%
Random mutants, on the other hand, serve as a way to \emph{selectively avoid
  exhaustiveness} when mutating values of large enumeration types --- or any
other type the user might want to use random mutations with.
%
Instead of mutating every numerical or character subexpression exhaustively, we
resolve in using a random generator and sample a small number of values from it
--- the precise amount is a tunable parameter of \mutagen.
%
Then, for instance, a mutator for integers becomes:

@@

\subsubsection{Mapping top-level mutations everywhere}

@@

Having selected the next test case, the main loop proceeds to execute it,
capturing both the result (valid, discarded due to preconditions, or failed)
and its execution trace over the system under test.
%
If the test case fails, it is immediately reported as a bug.
%
If not, the algorithm evaluates whether it was interesting (i.e., it exercises a
new path) based on its trace information and the one from previously executed
test cases (represented by \textit{TLog}).
%
If the test case does in fact discover a new path, it is enqueued on its
corresponding queue, depending on whether it was valid or got discarded.
%
%% (i.e., it failed the precondition).
%
This process alternates between generation and mutation until a bug is found or
we reach the test limit.

@@

\begin{algorithm}[t]
  \SetArgSty{textnormal}
  %% \SetInd{0em}{0.75em}
  \SetKw{KwNot}{not}
  \SetAlgoLined
  \DontPrintSemicolon
  \SetKwFunction{MG}{Loop}
  \SetKwProg{Fn}{Function}{:}{}
  \Fn{\MG{P, N, gen, mut}}{
    {\color{red}
    boring  $\gets$ 0\;
    reset   $\gets$ 1000\;
    R       $\gets$ 1\;
    }
    $\cdots$\;
    \While{i $<$ N}{
      {\color{red}
      \If{boring $>$ reset}{
        TLog   $\gets$ $\varnothing$\;
        reset $\gets$ reset * 2\;
        R     $\gets$ R * 2\;
      }
      }
      $\cdots$\;
      \lIf{\KwNot result}{\KwRet Bug(x)}
      \If{Interesting(TLog, trace)}{
        {\color{red} boring  $\gets$ 0}\;
        $\cdots$\;
      }
      \lElse {
        {\color{red} boring $\gets$ boring + 1}
      }
      $\cdots$\;
    }
  }

@@

We evaluated the performance of \mutagen using two case studies.
%
The first one is a simple abstract stack machine that enforces the
hyper-property \emph{noninterference} \cite{goguen1982security} using runtime
checks.
%
Its implementation was originally proven correct by
\citeauthor{10.1145/2578855.2535839} \citeyearpar{10.1145/2578855.2535839} in
Coq, and subsequently degraded by systematically introducing 20 bugs on its
enforcing mechanism.
%
\citeauthor{lampropoulos2019coverage} used this case study to compare \fuzzchick
against random testing approaches using na\"ive and manually-written smart
random generators.
%
In this work, we replicate their results and compare them against our tool.
%
Worth mentioning, since this case study was originally implemented in Coq, we
first translated it to Haskell in order to run \mutagen on its test suite.

The second case study aims to evaluate \mutagen in a realistic scenario,
and focuses on testing \textit{haskell-wasm} \cite{haskellwasm}, an existing
WebAssembly engine of industrial strength written in Haskell.
%
%
Sadly, comparing \mutagen with \fuzzchick on this case study has been out of the
scope of this work.
%
The reason for this is that porting \textit{haskell-wasm} into Coq, and later
adapting our test suite to work with \fuzzchick requires a subtantial effort.
%
Instead, we focus on comparing \mutagen against a more traditional RPBT approach
using \quickcheck, to evaluate its effectiveness versus the relative overhead of
the code instrumentation used in our tool.

@@

\noindent Control flow is achieved using the \texttt{Call} and \texttt{Ret}
instructions, which let the program jump back and forth within the instruction
memory using specially labeled values in the stack representing memory
addresses.
%
The argument in the \texttt{Push} instruction represents a value to be
inserted in the stack, whereas the argument of the \texttt{Call} instruction
encodes the number of elements of the stack to be treated as arguments.
%
Then, programs are simply modeled as sequences of instructions.
%
Finally, machine states can be modeled using a 4-tuple \textit{(pc, stk, m, i)}
that represents a particular configuration of the program counter, the stack and
the data- and instruction memories, respectively.
%
To preserve space, we encourage the reader to refer to the work of
\citeauthor{hritcu2013testing} \citeyearpar{hritcu2013testing,
  hrictcu2016testing} as well as the original CGPT paper
\cite{lampropoulos2019coverage} for more details about the implementation and
semantics of this case study.

\subsubsection{Single-Step Noninterference (SSNI)}

This abstract machine is designed to enforce noninterference, which is based on
the notion of \emph{indistinguishability}.
%
Intuitively, two machine states are indistinguishable from each other if they
only differ on secret data.
%
Using this notion, the particular variant of noninterference we are interested
in this work is called \emph{single-step noninterference}
\cite{hritcu2013testing} (SSNI).
%
SSNI asserts that, given two indistinguishable machine states, running a single
instruction on both machines brings them to resulting states that are also
indistinguishable.

@@

\subsubsection{IFC Rules}

The IFC rules enforced by this abstract machine are implemented using a rule
table indexed by the instruction that the machine is about to perform.
%
This table contains: the dynamic check that the abstract machine needs to
perform in order to enforce the IFC policy, along with the corresponding
security label of the program counter and the instruction result after the
operation is executed.
%
For instance, to execute the \texttt{Store} instruction (which stores a value in
a memory pointer) the machine checks that both the labels of the program counter
and the pointer together can flow to the label of the destination memory cell.
%
If this condition is not met, the machine then halts as indication of a
violation in the IFC policy.
%
After this check, this machine overwrites the value at the destination cell
and updates its label with the maximum sensibility of the involved labels.
%
In the rule table, this looks as follows:

\begin{center}
\scriptsize
\begin{tabular}{|c|c|c|c|}
\hline
\textbf{Instruction}
& \textbf{Precondition Check}
& \textbf{Final PC Label}
& \textbf{Final Result Label} \\
\hline
\texttt{Store}
& $l_{pc} \vee l_{p} \sqsubseteq l_{v}$
& $l_{pc}$
& $l_{v'} \vee l_{pc} \vee l_{p}$ \\
\hline
\end{tabular}
\end{center}

\noindent Where $l_{pc}$, $l_{p}$, $l_{v}$, and $l_{v'}$ represent the labels
of: the program counter, the memory pointer, and the old a new values stored in
that memory cell.
%
The symbol $\vee$ simply denotes the join of two labels, i.e., the
\emph{maximum} of their sensibilities.

Bugs can be systematically introduced in the IFC enforcing mechanism by removing
or weakening the checks stored in this rule table.

@@

WebAssembly \cite{haas2017bringing} is a popular assembly-like language designed
to be an open standard for executing low-level code in the web.%
%
\footnote{Although it has become increasingly popular in standalone, non-web
  contexts as well.}
%
WebAssembly programs are first validated (using a simple form of type-checking)
and later executed in a sandboxed environment (a virtual machine).
%
%% , making this language an attractive target for virtualization in
%% \emph{functions-as-a-service} platforms.
%
The language is relatively simple, in essence
%
\begin{inparaenum}
  \item it contains only four numerical types, representing both integers and
    IEEE754 floating-point numbers of either 32 or 64 bits;
  \item values of these types are by manipulated by functions written using
    sequences of stack instructions;
  \item functions are organized in modules and must be explicitly imported and
    exported;
  \item memory blocks can be imported, exported and grown dynamically;
\end{inparaenum}
%
among others.
%
%% To give the reader a taste, Fig. \ref{fig:wasm:factorial} the WebAssembly text
%% representation of module that exports a recursive implementation of the
%% factorial function.

WebAssembly semantics are fully specified, and programs are expected to be
consistently interpreted across engines --- despite some subtle details that we
will address soon.
%
Moreover, the WebAssembly standard provides a reference implementation with all
the functionality expected from a compliant engine.
%
%% \begin{wrapfigure}{r}{0.35\textwidth}
%% \vspace{-10pt}
%% \begin{verbatim}
%% (module
%% (func $f (param f64) (result f64)
%%   get_local 0
%%   f64.const 1
%%   f64.lt
%%   if (result f64)
%%     f64.const 1
%%   else
%%     get_local 0
%%     get_local 0
%%     f64.const 1
%%     f64.sub
%%     call $fact
%%     f64.mul
%%   end)
%% (export "fact" (func $f)))
%% \end{verbatim}
%% \caption{\label{fig:wasm:factorial}Factorial in WebAssembly}
%% \vspace{-10pt}
%% \end{wrapfigure}

Our tool is an attractive match for testing WebAssembly engines, and
\textit{haskell-wasm} in particular, as the space of WebAssembly programs that
can be represented using its AST contains mostly invalid ones, and automatically
derived random generators cannot satisfy all the invariants required to produce
interesting test cases.

In this work, we are interested in using \mutagen to test the two most complex
subsystems of \textit{haskell-wasm}: the \emph{validator} and the
\emph{interpreter} --- both being previously tested using a unit test suite.
%
We avoided spending countless hours writing an extensive property-based
specification to mimic the reference WebAssembly specification.
%
Instead, we used the reference WebAssembly implementation to find discrepancies
(that could potentially lead to bugs) via differential testing
\cite{mckeeman1998differential}.
%
In this light, our testing properties assert that any result produced by
\textit{haskell-wasm} matches that of the reference implementation.
%
Unsurprisingly, this engine had several latent bugs that were not caught by the
existing unit tests and that we discovered by \mutagen while developing our test
suite.
%
Moreover, \mutagen exposed two other discrepancies between \textit{haskell-wasm}
and the reference implementation.
%
Not severe enough to be classified as bugs, these discrepancies trigger parts of
the WebAssembly specification that are either not yet supported by the reference
implementation (multi-value blocks), or that produce a well-known
non-deterministic undefined behavior allowed by the specification (NaN
reinterpretation) \cite{perenyi2020stack}.
%
These findings are briefly outlined in Table \ref{table:wasm:bugs}.
%
All of them were reported and later confirmed by the authors of
\textit{haskell-wasm}.
%
Having sorted these issues out, we injected 10 new bugs in the validator as well
as 5 new bugs in the interpreter of this engine (see Table
\ref{table:wasm:injectedbugs}).
%
These bugs were mechanically injected by either
%
\begin{inparaenum}
\item removing an existing integrity check (to weaken the WebAssembly
  type-system/validator); or by
\item simulating a copy-and-paste induced bug.%
  \footnote{This kind of bugs was inspired by the real bug \#1 we found prior
    this step.}
  %
  For instance, by replacing the implementation of an operation with a similar
  compatible one (e.g., \texttt{i32.add} by \texttt{i32.sub}).
\end{inparaenum}

@@

\begin{table}[t]
\scriptsize
\begin{tabular}{|c|c|l|}
\hline
\textbf{Id}
& \textbf{Subsystem}
& \textbf{Description} \\
\hline
1
& Validator
& Wrong if-then-else type validation on else branch \\
\hline
2
& Validator
& Wrong stack type validation \\
\hline
3
& Validator
& Removed function type mismatch assertion \\
\hline
4
& Validator
& Removed max memory instances assertion \\
\hline
5
& Validator
& Removed function index out-of-range assertion \\
\hline
6
& Validator
& Wrong type validation on \texttt{i64.eqz} instruction \\
\hline
7
& Validator
& Wrong type validation on \texttt{i32} binary operations \\
\hline
8
& Validator
& Removed memory index out-of-range assertion \\
\hline
9
& Validator
& Wrong type validation on \texttt{i64} constants \\
\hline
10
& Validator
& Removed alignment validation on \texttt{i32.load} instruction \\
\hline
11
& Interpreter
& Wrong interpretation of \texttt{i32.sub} instruction \\
\hline
12
& Interpreter
& Wrong interpretation of \texttt{i32.lt\_u} instruction \\
\hline
13
& Interpreter
& Wrong interpretation of \texttt{i32.shr\_u} instruction \\
\hline
14
& Interpreter
& Wrong local variable initialization \\
\hline
15
& Interpreter
& Wrong memory address casting on \texttt{i32.load8\_s} instruction \\
\hline
\end{tabular}
\caption{\label{table:wasm:injectedbugs}Bugs injected into
  \textit{haskell-wasm}.}
%% \vspace{-15pt}
\end{table}

@@

\subsubsection{Testing the WebAssembly Validator}

%% We begin by designing a property to test the WebAssembly validator implemented
%% in \textit{haskell-wasm}.
%% %
To keep things simple, we assert that, whenever a randomly generated (or
mutated) WebAssembly module is valid according to \textit{haskell-wasm}, then
the reference implementation agrees upon it.
%
In other words, we are testing for false negatives.
%
In Haskell, we write the following property:

@@

Where the precondition (\texttt{isValidHaskellWasm m}) runs the input WebAssembly
module \texttt{m} against the \textit{haskell-wasm} validator, whereas the
postcondition (\texttt{isValidRefImpl m}) serializes \texttt{m} to a file, runs it
against the reference implementation validator and checks that no errors are
produced.

We note that, although we only focus on finding false negatives, in a complete
test suite, one would also want to test for false positives, i.e., when a module
is valid and \textit{haskell-wasm} rejects it.
%
%% This can be easily done by inverting the direction of the implication
%% (\texttt{<==}) in the property above.
%
However, the resulting property is much slower, as every test case will always
be serialized and run against the reference implementation.

@@

\subsubsection{Testing the WebAssembly Interpreter}

Testing the WebAssembly interpreter is substantially more complicated than
testing the validator, since it requires running and comparing the output of
actual programs.
%
To achieve this, we need the generated test cases to comply a with a common
interface that can be invoked both by \textit{haskell-wasm} and the reference
WebAssembly implementation.

To simplify things here as well, we write a function \texttt{mk\_module} to
build a stub module which initializes one memory block and exports a single
WebAssembly function.
%
\texttt{mk\_module} is parameterized by the definition of the single WebAssembly
function, along with its name and type signature.
%
%% In Haskell:
%
%% \begin{verbatim}
%% mk_module ty name fun =
%%   emptyModule { types     = [ ty ]
%%               , functions = [ fun ]
%%               , exports   = [ Export name (ExportFunc 0) ]
%%               , mems      = [ Memory (Limit 1 Nothing) ] }
%% \end{verbatim}
%
This way, we can define a testing property that can be instantiated with
randomly generated combinations of function types, function definitions and
lists of invocation arguments:

@@

This property instantiates the module stub using the input function and its type
signature, and uses it to invoke the function \texttt{foo} both on the
\textit{haskell-wasm} and reference implementation interpreter with the provided
arguments.
%
Then, the property asserts whether their results are equivalent.%
%
\footnote{In our implementation, we additionally set a short timeout to discard
  potentially diverging programs with infinite loops.}
%
Interestingly, equivalence in this context does not imply equality.
%
Non-deterministic operations in WebAssembly like NaN reinterpretations can
produce different equivalent results (as exposed by the discrepancy \#5 in Table
\ref{table:wasm:bugs}), and our equivalence relation needs to take that into
account.
%

@@

This specialized property lets us generate functions using this fixed type and
invoke them with the exact number and type of arguments required.
%
In our experiments (presented in the next section), we use this property when
finding all the injected bugs into the \textit{haskell-wasm} interpreter.
%
As before, we remark that a complete test suite should at least include
different variants of this property testing functions of several different
types, as well as properties testing multiple functions simultaneously.

@@

We performed all experiments in a dedicated workstation with an Intel Core
i7-8700 CPU running at 3.20GHz, and equipped with 32GB of RAM.
%
Each experiment was repeated 30 times except for the ones testing the
WebAssembly interpreter, which were run 10 times due to time constraints.
%
From there, we followed the approach taken by
\citeauthor{lampropoulos2019coverage} and collected the mean-time-to-failure
(MTTF) of each bug, i.e., how quickly a bug can be found in wall clock time.
%
We used a one-hour timeout to stop the execution of both tools if they have not
yet found a counterexample.
%
In addition, we collected the failure rate observed for each bug, i.e. the
proportion of times each tool finds each bug within the one-hour testing budget.
%
%% We found this metric crucial to be analyzed when replicating \fuzzchick's
%% results, as opposed to just paying attention at the MTTF.

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=\linewidth]{tikz/ifc-mttf.tikz}
  \includegraphics[width=\linewidth]{tikz/ifc-fr.tikz}
  %% \vspace{-20pt}
  \caption{\label{fig:results:ifc} Comparison between \fuzzchick and \mutagen
    across 20 different bugs for the IFC stack machine. }
%% \vspace{-10pt}
\end{figure}

The results of this case study are shown in Fig. \ref{fig:results:ifc} --- we
encourage the reader to obtain a color copy of this work.
%
The injected bugs are ordered by the failure rate achieved by \fuzzchick in
decreasing order.
%
Moreover, notice the logarithmic scale used on the MTTF.

As introduced earlier, \fuzzchick can only find 5 out of the 20 bugs injected
with a 100\% success rate within the one hour testing budget.
%
In turn, our tool manages to find every bug on all runs, and taking less than a
minute in the worst absolute case.
%
These results are somewhat pessimistic towards \fuzzchick if compared to the
ones presented originally by \citeauthor{lampropoulos2019coverage}.
%
However, we remark that our experiments encompass 30 independent runs instead of
the 5 originally used when presenting \fuzzchick
\citeyearpar{lampropoulos2019coverage}.
%
Moreover, since the MTTF simply aggregates all runs, regardless of if they found
a bug or timed out, this metric is quite sensitive to the timeout used on each
experiment, and will inflate the results as soon as the failure rate goes below
1.
%
For this reason, comparing the failure rate between tools give us a better
estimation of the reliability of both tools, where \mutagen shows a clear
improvement.

>>>>

%
% FuzzChick Limitations
%
\paragraph{Limitations of CGPT}

@@

%% At this point, the reader might consider:
%% %
%% what about large enumeration types like integers or characters?
%% %
%% Does \mutagen transform them exaustively too?
%% %
%% \emph{Certainly not.}
%
Random mutants, on the other hand, serve as a way to \emph{selectively avoid
  exhaustiveness} when mutating values of large enumeration types --- or any
other type the user might want to use random mutations with.
%
Instead of mutating every numerical or character subexpression exhaustively, we
sample a small number of values from a given random generator --- this amount is
a tunable parameter of \mutagen.
%
Then, for instance, a mutator for integers becomes:

@@

\paragraph{Mapping top-level mutations everywhere}

@@

Having selected the next test case, the main loop proceeds to execute it,
capturing both the result (valid, discarded, or failed)
and its execution trace over the system under test.
%
If the test case fails, it is reported as a bug.
%
If not, the algorithm evaluates whether it was interesting (i.e., it exercises a
new path) based on its trace information and the one from previously executed
test cases (represented by \textit{TLog}).
%
If the test case discovers a new path, it is enqueued on its corresponding
queue.
% depending on whether it was valid or got discarded.
%
%% (i.e., it failed the precondition).
%
This process alternates between generation and mutation until it finds a bug or
reaches the test limit N.

@@

\begin{algorithm}[t]
  \SetArgSty{textnormal}
  %% \SetInd{0em}{0.75em}
  \SetKw{KwNot}{not}
  \SetAlgoLined
  \DontPrintSemicolon
  \SetKwFunction{MG}{Loop}
  \SetKwProg{Fn}{Function}{:}{}
  \Fn{\MG{P, N, gen, mut}}{
    {\color{red} boring $\gets$ 0}; {\color{red} reset $\gets$ 1000}; {\color{red}R $\gets$ 1}\;
    $\cdots$\;
    \While{i $<$ N}{
      {\color{red}
      \If{boring $>$ reset}{
        TLog   $\gets$ $\varnothing$\;
        reset $\gets$ reset * 2\;
        R     $\gets$ R * 2\;
      }
      }
      $\cdots$\;
      \lIf{\KwNot result}{\KwRet Bug(x)}
      \If{Interesting(TLog, trace)}{
        {\color{red} boring  $\gets$ 0}\;
        $\cdots$\;
      }
      \lElse {
        {\color{red} boring $\gets$ boring + 1}
      }
      $\cdots$\;
    }
  }

@@

We evaluated the performance of \mutagen using two case studies.
%
The first one is a simple abstract stack machine that enforces the
\emph{noninterference} \cite{goguen1982security} using runtime checks.
%
Its implementation was proven correct by \citeauthor{10.1145/2578855.2535839}
\citeyearpar{10.1145/2578855.2535839} in Coq, and subsequently degraded by
systematically introducing 20 bugs on its enforcing mechanism.
%
\citeauthor{lampropoulos2019coverage} used this case study to compare \fuzzchick
against RPBT using na\"ive- and hand-tuned random generators.
%
Here, we replicate their results and compare them against our tool.
%
Worth mentioning, we translated this case study into Haskell in order to run
\mutagen on its test suite.

The second case study aims to evaluate \mutagen in a realistic scenario, and
targets \textit{haskell-wasm} \cite{haskellwasm}, an existing WebAssembly engine
of industrial strength written in Haskell.
%
%
Sadly, comparing \mutagen with \fuzzchick on this case study has been out of the
scope of this work, since porting \textit{haskell-wasm} into Coq, and later
adapting our test suite to work with \fuzzchick requires a subtantial effort.
%
Instead, we compare \mutagen against a traditional RPBT approach, evaluating its
effectiveness versus the relative overhead of the code instrumentation used by
our tool.

@@

\noindent Control flow is achieved using the \texttt{Call} and \texttt{Ret}
instructions, which let the program jump back and forth within the instruction
memory using specially labeled values in the stack representing memory
addresses.
%
The argument in the \texttt{Push} instruction represents a value to be
inserted in the stack, whereas the argument of the \texttt{Call} instruction
encodes the number of elements of the stack to be treated as arguments.
%
Then, programs are simply modeled as sequences of instructions.
%
Finally, machine states are modeled using a 4-tuple \textit{(pc, stk, m, i)}
representing a particular configuration of the program counter, the stack, and
the data- and instruction memories, respectively.
%
To preserve space, we encourage the reader to refer to the work of
\citeauthor{hritcu2013testing} \citeyearpar{hritcu2013testing,
  hrictcu2016testing} and \citeauthor{lampropoulos2019coverage} for more details
about the implementation and semantics of this case study.

\paragraph{Single-Step Noninterference (SSNI)}

This abstract machine is designed to enforce noninterference, which is based on
the notion of \emph{indistinguishability}.
%
Intuitively, two machine states are indistinguishable if they only differ on
secret data.
%
Using this notion, the variant of noninterference we are interested in is called
\emph{single-step noninterference} \cite{hritcu2013testing} (SSNI).
%
Given two indistinguishable machine states, SSNI asserts that running a single
instruction on both machines brings them to resulting states that are also
indistinguishable.

@@

\paragraph{IFC Rules}

The IFC rules enforced by this machine are encoded in a table indexed by its
different instructions.
%
This table contains: the dynamic check that the machine needs to perform to
enforce the IFC policy, along with the corresponding security label of the
program counter and the instruction result after the instruction is executed.
%
For instance, to execute the \texttt{Store} instruction (which stores a value in
a memory pointer) the machine checks that both the labels of the program counter
and the pointer together can flow to the label of the destination memory cell.
%
If this condition is not met, the machine then halts as indication of a
violation in the IFC policy.
%
After this check, the machine overwrites the value at the destination cell and
updates its label with the maximum sensibility of the involved labels.
%
In the rule table, this looks as follows:

%% \vspace{-10pt}
\begin{center}
\scriptsize
\begin{tabular}{|c|c|c|c|}
\hline
\textbf{Instruction}
& \textbf{Precondition Check}
& \textbf{Final PC Label}
& \textbf{Final Result Label} \\
\hline
\texttt{Store}
& $l_{pc} \vee l_{p} \sqsubseteq l_{v}$
& $l_{pc}$
& $l_{v'} \vee l_{pc} \vee l_{p}$ \\
\hline
\end{tabular}
\end{center}

\noindent Where $l_{pc}$, $l_{p}$, $l_{v}$, and $l_{v'}$ represent the labels
of: the program counter, the memory pointer, and the old a new values stored in
that memory cell.
%
The symbol $\vee$ simply denotes the join of two labels, i.e., the
\emph{maximum} of their sensibilities.
%
Later, bugs are systematically introduced in the IFC enforcing mechanism by
removing or weakening the checks stored in this rule table.

@@

WebAssembly \cite{haas2017bringing} is an assembly-like language designed for
executing low-level code in the web.
%% %
%% \footnote{Although it has become increasingly popular in standalone, non-web
%%   contexts as well.}
%
WebAssembly programs are first validated
%
%(using a simple form of type-checking)
%
and later executed in a sandboxed environment.
%
%% , making this language an attractive target for virtualization in
%% \emph{functions-as-a-service} platforms.
%
The language is relatively simple, in essence
%
\begin{inparaenum}
  \item it contains only four numerical types, representing both integers and
    IEEE754 floating-point numbers of either 32 or 64 bits;
  \item values of these types are manipulated by functions written using
    sequences of stack instructions;
  \item functions are organized in modules and must be explicitly imported and
    exported;
  \item memory blocks can be imported, exported and grown dynamically;
\end{inparaenum}
%
among others.
%
%% To give the reader a taste, Fig. \ref{fig:wasm:factorial} the WebAssembly text
%% representation of module that exports a recursive implementation of the
%% factorial function.
%
WebAssembly semantics are fully specified, and programs must be consistently
interpreted across engines --- despite some subtle details that we will address
soon.
%
For this, the WebAssembly standard provides a reference implementation with all
the functionality expected from a compliant engine.
%
%% \begin{wrapfigure}{r}{0.35\textwidth}
%% \vspace{-10pt}
%% \begin{verbatim}
%% (module
%% (func $f (param f64) (result f64)
%%   get_local 0
%%   f64.const 1
%%   f64.lt
%%   if (result f64)
%%     f64.const 1
%%   else
%%     get_local 0
%%     get_local 0
%%     f64.const 1
%%     f64.sub
%%     call $fact
%%     f64.mul
%%   end)
%% (export "fact" (func $f)))
%% \end{verbatim}
%% \caption{\label{fig:wasm:factorial}Factorial in WebAssembly}
%% \vspace{-10pt}
%% \end{wrapfigure}

\begin{table}[t]
\scriptsize
\begin{tabular}{|c|c|l|}
\hline
\textbf{Id}
& \textbf{Subsystem}
& \textbf{Description} \\
\hline
1
& Validator
& Wrong if-then-else type validation on else branch \\
\hline
2
& Validator
& Wrong stack type validation \\
\hline
3
& Validator
& Removed function type mismatch assertion \\
\hline
4
& Validator
& Removed max memory instances assertion \\
\hline
5
& Validator
& Removed function index out-of-range assertion \\
\hline
6
& Validator
& Wrong type validation on \texttt{i64.eqz} instruction \\
\hline
7
& Validator
& Wrong type validation on \texttt{i32} binary operations \\
\hline
8
& Validator
& Removed memory index out-of-range assertion \\
\hline
9
& Validator
& Wrong type validation on \texttt{i64} constants \\
\hline
10
& Validator
& Removed alignment validation on \texttt{i32.load} instruction \\
\hline
11
& Interpreter
& Wrong interpretation of \texttt{i32.sub} instruction \\
\hline
12
& Interpreter
& Wrong interpretation of \texttt{i32.lt\_u} instruction \\
\hline
13
& Interpreter
& Wrong interpretation of \texttt{i32.shr\_u} instruction \\
\hline
14
& Interpreter
& Wrong local variable initialization \\
\hline
15
& Interpreter
& Wrong memory address casting on \texttt{i32.load8\_s} instruction \\
\hline
\end{tabular}
\caption{\label{table:wasm:injectedbugs}Bugs injected into
  \textit{haskell-wasm}.}
\vspace{-10pt}
\end{table}

Our tool is an attractive match for testing WebAssembly engines:
%
%% and \textit{haskell-wasm} in particular,
%
most of the programs that can be represented using WebAssembly's AST are
invalid, and automatically derived random generators cannot satisfy the
invariants required to produce interesting test cases.

In this work, we are interested in using \mutagen to test the two most complex
subsystems of \textit{haskell-wasm}: the \emph{validator} and the
\emph{interpreter} --- both being previously tested using a unit test suite.

We avoided spending countless hours writing an extensive property-based
specification to mimic the WebAssembly specification.
%
Instead, we used the reference implementation to find discrepancies (that could
potentially lead to bugs) via differential testing
\cite{mckeeman1998differential}.
%
In this light, our testing properties assert that any result produced by
\textit{haskell-wasm} matches that of the reference implementation.
%
Notably, this engine had several latent bugs that were not caught by the
existing unit tests and that we discovered by \mutagen while developing our test
suite.
%
Moreover, \mutagen exposed two other discrepancies between \textit{haskell-wasm}
and the reference implementation.
%
These discrepancies trigger parts of the specification that are either not yet
supported by the reference implementation (multi-value blocks), or that produce
a well-known non-deterministic undefined behavior (NaN reinterpretation)
\cite{perenyi2020stack}.
%
These findings are briefly outlined in Table \ref{table:wasm:bugs}.
%
All of them were reported and later confirmed by the authors of
\textit{haskell-wasm}.
%
Having sorted these issues out, we injected 10 new bugs in the validator as well
as 5 new bugs in the interpreter of this engine (see Table
\ref{table:wasm:injectedbugs}).
%
These bugs were mechanically injected by either
%
\begin{inparaenum}
\item removing an existing integrity check (to weaken the WebAssembly
  type-system/validator); or by
\item simulating a copy-and-paste induced bug,%
  \footnote{This kind of bugs was inspired by the real bug \#1 we found prior
    this step.}
  %
  replacing the implementation of an instruction with a compatible one
  (e.g., \texttt{i32.add} by \texttt{i32.sub}).
\end{inparaenum}

@@

@@

\paragraph{Testing the WebAssembly Validator}

%% We begin by designing a property to test the WebAssembly validator implemented
%% in \textit{haskell-wasm}.
%% %
To keep things simple, we assert that, whenever a randomly generated (or
mutated) WebAssembly module is valid according to \textit{haskell-wasm}, then
the reference implementation agrees upon it.
%
%% In other words, we are testing for false negatives.
%
In Haskell, we write the property:

@@

\noindent The precondition (\texttt{isValidHaskellWasm m}) runs the input
WebAssembly module \texttt{m} against \textit{haskell-wasm}'s validator, whereas
the postcondition (\texttt{isValidRefImpl m}) serializes \texttt{m}, runs it
against the reference implementation and checks that no errors are produced.

We note that, although we only focus on finding false negatives, a comprehensive
test suite should also test for false positives, i.e., when a module is valid
and \textit{haskell-wasm} rejects it.
%
%% This can be easily done by inverting the direction of the implication
%% (\texttt{<==}) in the property above.
%
%% However, the resulting property is much slower, as every test case will always
%% be serialized and run against the reference implementation.

@@

\paragraph{Testing the WebAssembly Interpreter}

Testing the WebAssembly interpreter is substantially more complicated than
testing the validator, since it requires actually running and comparing the
output of test case programs.
%
To achieve this, the generated test cases need to comply a with a common
interface that can be invoked both by \textit{haskell-wasm} and the reference
implementation.
%
To simplify things, we write a function \texttt{mk\_module} to build a stub
module which initializes one memory block and exports a single WebAssembly
function.
%
Such function is parameterized by the definition of the single function, along
with its name and type signature.
%
%% In Haskell:
%
%% \begin{verbatim}
%% mk_module ty name fun =
%%   emptyModule { types     = [ ty ]
%%               , functions = [ fun ]
%%               , exports   = [ Export name (ExportFunc 0) ]
%%               , mems      = [ Memory (Limit 1 Nothing) ] }
%% \end{verbatim}
%
Then, we can use \texttt{mk\_module} to define a testing property parameterized
by a function type, along with its definition and invocation arguments:

%% define a testing property parameterized by a function type, a
%% function body and invocation arguments:

@@

This property instantiates a module stub \texttt{m} using the input function
(\texttt{fun}) and its type signature (\texttt{ty}).
%
Then, it invokes the function \texttt{foo} of the module \texttt{m} both on
\textit{haskell-wasm} and the reference interpreter with the provided arguments
(\texttt{args}).
%
Then, the property asserts whether their results are equivalent.%
%
\footnote{In our implementation, we additionally set a short timeout to discard
  potentially diverging programs with infinite loops.}
%
Interestingly, equivalence in does not imply equality.
%
Non-deterministic operations in WebAssembly like NaN reinterpretations can
produce different equivalent results (as exposed by the discrepancy \#5 in Table
\ref{table:wasm:bugs}), and our equivalence relation needs to take that into
account.
%

@@

\noindent This property lets us generate functions with a fixed type and invoke
them with the exact number and type of arguments required.
%
%% In our experiments (presented in \S \ref{sec:evaluation}),
%
We use this property when finding all the injected bugs into the
\textit{haskell-wasm} interpreter in the next section.
%
%% As before, we remark that a complete test suite should at least include
%% different variants of this property testing functions of several different
%% types, as well as properties testing multiple functions simultaneously.

@@

We performed all experiments in a dedicated workstation with 32GB of RAM and an
Intel Core i7-8700 CPU running at 3.20GHz.
%
We repeated each experiment 30 times except for the ones testing the WebAssembly
interpreter, which were run 10 times due to time constraints.
%
From there, we followed the approach taken by
\citeauthor{lampropoulos2019coverage} and collected the mean-time-to-failure
(MTTF) of each bug, i.e., how quickly a bug can be found in wall clock time.
%
We used a one-hour timeout to stop the execution of both tools if they have not
yet found a counterexample.
%
In addition, we collected the failure rate (FR) observed for each bug, i.e. the
proportion of times each tool finds each bug within the one-hour testing budget.
%
%% We found this metric crucial to be analyzed when replicating \fuzzchick's
%% results, as opposed to just paying attention at the MTTF.

@@

\begin{figure}[b]
  \centering
  \includegraphics[width=\linewidth]{tikz/ifc-mttf.tikz}
  \includegraphics[width=\linewidth]{tikz/ifc-fr.tikz}
  \vspace{-20pt}
  \caption{\label{fig:results:ifc} Comparison between \fuzzchick and \mutagen
    across 20 different bugs for the IFC stack machine. }
\vspace{-5pt}
\end{figure}

The results of this case study are shown in Fig. \ref{fig:results:ifc}.
%
%% --- we encourage the reader to obtain a color copy of this work.
%
The injected bugs are ordered by the failure rate achieved by \fuzzchick in
decreasing order.
%
Moreover, notice the logarithmic scale used on the MTTF.

As introduced earlier, \fuzzchick can only find 5 out of the 20 bugs injected
with a 100\% success rate within the one hour testing budget.
%
In turn, our tool manages to find every bug on all runs, and taking less than a
minute in the worst absolute case.
%
While these results are somewhat pessimistic towards \fuzzchick if compared to
the ones presented originally by \citeauthor{lampropoulos2019coverage}, we
remark that our experiments encompass 30 independent runs instead of the 5 used
in the original study.
%
Moreover, since the MTTF simply aggregates all runs, regardless of if they found
a bug or timed out, this metric is sensitive to the timeout used on each
experiment, and will inflate the results as soon as the failure rate goes below
1.
%
For this reason, comparing the failure rate between tools gives a better
estimation of their reliability, where \mutagen shows a clear improvement.

%% If we only consider the succesfull runs where \fuzzchick does not time out, we
After carefully analyzing the results obtained using \fuzzchick, we observed a
peculiarity.
%
This tool either finds bugs quickly (after a few thousands tests) or does not
find them at all within the one-hour budget.
%% , suggesting that its power
%% scheduler assigns some enqueued test cases too little energy before they become
%% uninteresting and get ultimately thrown away.
%
With this in mind, it is fair to think that \fuzzchick might be a better choice
if we set a shorter timeout and run it several times.
%
Thus, to be an improvement over \fuzzchick, our tool must find bugs not only
more reliably, but also fast!

\setlength{\intextsep}{10pt}%
\setlength{\columnsep}{10pt}%
\begin{wrapfigure}{r}{0.5\linewidth}
%% \begin{wrapfigure}{r}{0.5\textwidth}
%% \centering
\vspace{-10pt}
\scriptsize
